---
title: "Bootstrap"
author: "Marcin Mazurek"
date:  '2017-11-30'
output:
  html_document:
    toc: yes
  html_notebook:
    toc: yes
  pdf_document:
    toc: yes
---




# Resampling

A method for estimating prediction error.

## Test set extracted from a simple sample:
1. Classical approach: If we have a sufficiently large sample, we extract a validation (test) set from it.

```{r}
library(ISLR)
head(Auto)

train.idx <- sample(1:nrow(Auto), round(0.7*nrow(Auto)))
train<-Auto[train.idx,]
test<-Auto[-train.idx,]

lmfit<-glm(mpg~horsepower+cylinders+ weight, data=train)
summary(lmfit)

mse=mean((test$mpg - predict(lmfit, test))^2)

# Estimation of prediction error based on the test set:
mse
```


## Cross Validation
The dataset is divided into ${k}$ partitions.

K-fold: one of the ${k}$ partitions is used for testing, the remaining ${k-1}$ partitions are used for model building. The procedure is repeated ${k}$ times (for each partition used as the test set).

LOOCV: a special case of K-fold where the number of partitions equals the number of observations.

### Leave-One-Out


```{r}
se_total=0
for (i in 1:nrow(Auto)){
  train = Auto[-i,]
  test = Auto[i,] 
  lmfit<-glm(mpg~horsepower+cylinders+ weight, data=train)
  se=(test$mpg - predict(lmfit, test))^2
  se_total = se_total + se 
}
 mse = se_total / nrow(Auto)
 mse 
```

The same operation using the boot library:


```{r}
library(ISLR)
library(boot)
lmfit<-glm(mpg~horsepower+cylinders+ weight, data=Auto)

# LOOCV:
k=nrow(Auto)
m<-cv.glm(data=Auto, lmfit, K=k)

# Error:
mse = m$delta[1]
mse 
```



### k-fold Cross validation



```{r}

lmfit<-glm(mpg~horsepower+cylinders+ weight, data=Auto)

# LOOCV:
m<-cv.glm(data=Auto, lmfit, K=5)

# Error calculated based on the average of 5 folds:
mse = m$delta[1]
mse 




```




## Bootstrap

Pull oneself up by one's bootstrap - to get out of trouble using one's own resources

A method for estimating the distribution of estimators when we do not make any assumptions about the distribution of variables in the population.

1. We do not know the distribution F
2. We only have a single realization of a simple random sample from this distribution

### Estimating the distribution of a statistic using the bootstrap method



General outline of the procedure:

Sampling distribution of statistic S:

$S=S(x_1, x_2, \cdots, x_n)$

based on a simple random sample:

$\mathbf{X}=(X_1, X_2, \cdots, X_n)$

1. Create samples by sampling with replacement: $\mathbf{x_1^{*}}, \cdots, \mathbf{x_M^{*}}$ : bootstrap sample

2. Calculate the statistic of interest $S(\mathbf{x_1^{*}}), \cdots, S(\mathbf{x_M^{*}})$ for each sample. The distribution of the statistic obtained from the samples is the bootstrap distribution.

3. The bootstrap distribution provides information about the distribution of the statistic S




### Bootstrap - estimating distributions of estimators

```{r}
library(ISLR)
head(Auto)
lmfit<-glm(mpg~horsepower+cylinders+ weight, data=Auto)
summary(lmfit)
coef<-matrix(nrow=30, ncol=4)
nrow(Auto)
for (i in 1:30){
  index <- sample(1:nrow(Auto), nrow(Auto), replace=TRUE) 
  train <- Auto[index,]
  test  <- Auto[-index,]
  # training the network, linear.output = FALSE because variables are binary
  lmfit<-glm(mpg~horsepower+cylinders+ weight, data=train)
  coef[i,]<-lmfit$coefficients
}

coefficients<-colMeans(coef)
coef_std <- apply(coef, MARGIN =2, FUN=sd )

coefficients
coef_std


hist(coef[,1])
head(Auto)


train<-Auto
lmfit<-glm(mpg~horsepower+cylinders+ weight, data=train)
summary(lmfit)
```




### Classification using neural networks
```{r}
library(neuralnet)
library(nnet)
data(iris)


# one-hot encoder - binary encoding of the variable
iris$setosa<-  ifelse(iris$Species == "setosa",1,0)
iris$versicolor<-ifelse(iris$Species == "versicolor",1,0)
iris$virginica<-ifelse(iris$Species == "virginica",1,0)

# instead of this code, you can use from the nnet library:
species_enc<-class.ind(iris$Species)


# normalize input attributes and attach three binary output columns
iris<-cbind(scale(iris[,1:4]), iris[,6:8])


# example of building a neural network with one hidden layer
f<-as.formula("setosa+versicolor+virginica~Sepal.Width + Sepal.Length + Petal.Width + Petal.Length")
nn<-neuralnet(f, hidden=c(1), data=iris, linear.output=FALSE)
plot(nn)
pred<-compute(nn,iris[,1:4])$net.result
pred_idx<-max.col(pred)
actuals<-iris[,5:7] 
actuals_idx<-max.col(actuals)
# percentage model accuracy: TRUE(1) for matching records, FALSE(0) for errors
accuracy<- mean(actuals_idx==pred_idx)
accuracy

```


### Bootstrap 0.632 in classification

$n$ - number of observations in the data sample, we sample with replacement $n$ observations.

The probability of not drawing an observation is:
$(1- \frac{1}{n})^n \approx e^{-1} \approx 0.368$,

so the expected number of unique observations in the drawn set is $0.632n$.

Accuracy estimation is performed with $b$ bootstrap samples, let $acc_{i}$ be the estimate of classification accuracy for the ${i}$-th bootstrap sample.

$acc_{boot} = \frac{1}{b} \sum_{i=1}^{b}(0.632*acc_{i}+0.368*acc_i)$

Variance can be estimated based on the variance of the estimator


```{r}
# example of 0.632 bootstrap for estimating classification error
acc_table = NULL 

for (i in 1:10){
  index <- sample(1:nrow(iris), nrow(iris), replace=TRUE) 
  train <- iris[index,]
  test  <- iris[-index,]
  # training the network, linear.output = FALSE because variables are binary
  nn<-neuralnet(f, hidden=c(1),  data=iris, linear.output = FALSE)
  # calculate predictions
  pred<-compute(nn,test[,1:4])$net.result
  pred_idx<-max.col(pred)
  # actuals
  actuals<-test[,5:7] 
  actuals_idx<-max.col(actuals)
  
  # proportion of correctly classified:
  accuracy<- mean(actuals_idx==pred_idx)
  acc_table[i]= accuracy
}

acc_table

# calculation of prediction error


avg_acc<- mean(acc_table)
avg_acc

acc_boot<-mean(0.632*acc_table+0.368*accuracy)
acc_boot


```



